---
title: "Data Manipulation with R"
author: "Wenxiao Zhou"
date: "4/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Manipulation with dplyr

**Transforming Data with dplyr**\
&NewLine;
```{r}
#Understanding the data
library(dplyr)
help(glimpse) #within dplyr package
setwd("/Users/zhouwenxiao/Desktop/Data Manipulation with R")
counties<-readRDS("counties.rds")
#take a look at data using glimpse() function
glimpse(counties)
```
```{r}
#select columns
#select 'state','county','population','poverty' columns
counties %>% select(state,county,population,poverty)
```

The `arrange()` verb sorts your data based on one or more variables. You can start with your counties selected variable, then the `pipe operator`, then `arrange()`. Inside the parentheses, put population to specify the variable you want to sort by. 
```{r}
#arrange observations
#sort the selected table using `public_work` variable in descending order
counties_selected <- counties %>%
  select(state, county, population, private_work, public_work, self_employed)
counties_selected %>%
  arrange(desc(public_work))
```

You can use the `filter()` verb to extract only particular observations from a dataset, based on a condition. Recall that after your first verb, you can add a pipe operator, then add another verb. You can pipe any number of verbs together to transform your dataset. For example, after the `arrange()`, you could add filter state equals equals quote New York to get only counties in the state of New York. Notice that the observations are filtered, but they're still sorted by population thanks to your arrange.

```{r}
#filter for conditions
#find only the counties that have a population above one million(1000000)
#and also only state of California
counties_selected %>%
  filter(population>1000000,state=='California')
#Finally, you can combine multiple conditions together in a filter
```
```{r}
#filter & arrange together
counties_selected %>%
  # Filter for Texas and more than 10000 people
  filter(state=="Texas", population>10000) %>%
  # Sort in descending order of private_work
  arrange(desc(private_work))
```

#Use `mutate()` verb to add new variables or change existing variables.
```{r}
#calculate the number of government employees
#within the counties data frame, unemployment is a percentage variable, same 
#as public_work variable
counties_selected<-counties %>%
    select(state, county, population, public_work)
counties_selected %>% 
    # add a new column public_workers with the number of people employed in public work
   mutate(public_workers=population*public_work/100) %>%
  # sort in descending order of the public_workers column
   arrange(desc(public_workers))
```
```{r}
#calculate the percentage of women in a county
counties_selected <- counties %>%
  # Select the columns state, county, population, men, and women
  select(state,county,population,men,women)
counties_selected %>%
   mutate(proportion_women=women/population)
```

Now, we can combine all four verbs together 
`select()`,`mutate()`,`filter()`,`arrange()`
```{r}
counties %>%
  # Select the five columns 
   select(state,county,population,men,women) %>%
  # Add the proportion_men variable
  mutate(proportion_men=men/population) %>%
  # Filter for population of at least 10,000
  filter(population>=10000) %>%
  # Arrange proportion of men in descending order
  arrange(desc(proportion_men))
```



**Aggregating Data**\
&NewLine;
To `filter` or `sort` for observations you're interested in, and to `add new variables with the mutate verb` are the work at the same level as the initial data, where every observation corresponds to one US county.\
Now you need more about the summary, which is called as aggregate data: to take many observations and summarize them into one. This is a common data science strategy for making data manageable and interpreted.

```{r}
#counting by region
#Finding the number of counties in each region
counties_selected <- counties %>%
  select(county, region, state, population, citizens)
counties %>%
   count()
counties %>% 
  count(state)
counties_selected %>%
  count(region,sort=TRUE)
```
```{r}
#Count the number of counties in each state, weighted based on the `citizens` column, and sorted in descending order
counties_selected <- counties %>%
  select(county, region, state, population, citizens) %>%
  count(state,wt=citizens,sort=TRUE)
counties_selected
```

Mutating and counting\
Use multiple verbs together to answer increasing complicated questions of your data.
```{r}
#walk: which offers a percentage of people in each county that walk to work, to add a new column and count based on it
counties_selected <- counties %>%
  select(county, region, state, population, walk)
counties_selected %>%
  # Add population_walk containing the total number of people who walk to work 
  mutate(population_walk = population * walk / 100) %>%
  # Count weighted by the new column, sort in descending order
  count(state, wt = population_walk, sort = TRUE)
```

`count()` is a special case of a more general set of verbs: `group_by() and summarize()`.\
The `summarize()` verb takes many observations and turns them into one observation.\
&Newline;
&Newline;

```{r}
#summarizing
counties_selected <- counties %>%
  select(county, population, income, unemployment)
#summarize the counties dataset to find the following columns:
#min_population (smallest population)
#max_unemployment (maximum unemployment)
#average_income (mean of the income variable)
counties_selected %>%
  # Summarize to find minimum population, maximum unemployment, and average income
  summarize(min_population=min(population),
            max_unemployment=max(unemployment),
            average_income=mean(income))
```
```{r}
#summarize by state
#land_area: shows the land area in square miles
#summarize both population and land area by state, and
#find the density (in people per square miles)
counties_selected <- counties %>%
  select(state,county, population, land_area)
counties_selected %>%
  group_by(state) %>%
  summarize(total_area = sum(land_area),
            total_population = sum(population)) %>%
  # Add a density column (people per square mile)
  mutate(density=total_population/total_area) %>%
  # Sort by density in descending order
  arrange(desc(density))
```
```{r}
#summarize by state and region
counties_selected <- counties %>%
  select(region, state, county, population)
counties_selected %>%
  # Group and summarize to find the total population
  group_by(region, state) %>%
  summarize(total_pop = sum(population)) %>%
# Calculate the average_pop and median_pop columns 
  # Notice that the tibble is still grouped by region, use another 
  # summarize() step to add more columns under this grouping

  #the average state population in each region
  #the median state population in each region 
  summarize(average_pop=mean(total_pop),
            median_pop=median(total_pop))
#The final result includes four regions' average population and median population by two columns
```


`top_n()` is helpful to keep the most extreme observations from each group\
Just like `summarize()`, top_n() operates on  a grouped table. The function takes two arguments: the number of observations you want from each group\
               the column you want to weight by\
`top_n()` is often used when creating graphs, where we're interested in pulling the extreme examples to include in the visualization.               
```{r}
#selecting a county from each region
counties_selected <- counties %>%
  select(region, state, county, metro, population, walk)
#Find the county in each region with the highest percentage of citizens who walk to work
counties_selected %>%
  group_by(region) %>%
  top_n(1,walk)
```
```{r}
#finding the highest income state in each region
#group_by(), summarize(), top_n()
counties_selected <- counties %>%
  select(region, state, county, population, income)
counties_selected %>%
  group_by(region, state) %>%
  # Calculate average income
  summarize(average_income=mean(income)) %>%
  # Find the highest income state in each region
  top_n(1,average_income)
# From our results, we can see that the New Jersey in the Northeast is the state with the highest average_income of 73014.
```

When you group by multiple columns and then summarize, it's important to remember that the summarize "peels off" one of the groups, but leaves the rest on.\
For example, if you group_by(X, Y) then summarize, the result will still be grouped by X
```{r}
#Using summarize, top_n,count together
#Answer the following questions:
# In How many states do more people live in metro areas than non-metro areas?
counties_selected <- counties %>%
  select(state, metro, population)
counties_selected %>%
  # Find the total population for each combination of state and metro
  group_by(state, metro) %>%
  summarize(total_pop = sum(population)) %>%
  # Extract the most populated row for each state
  top_n(1, total_pop) %>%
  # Count the states with more people in Metro or Nonmetro areas
  ungroup()   %>%
  # (firstly, since most data operations are done on groups defined by variables. group_by() takes an existing table and converts it into a grouped tbl where operations are performed "by group". ungroup() removes grouping.)
  count(metro)
```


**Selecting and Transforming Data**\
```{r}
#using a colon (:) can select variables that are consecutive
counties %>%
  # Select state, county, population, and industry-related columns
  select(state,county,population,professional:production) %>%
  # Arrange service in descending order 
  arrange(desc(service))
```
```{r}
#use starts_with(), ends_with(), which finds the columns that end with a particular string
counties %>%
  # Select the state, county, population, and those ending with "work"
  select(state,county,population,ends_with("work")) %>%
  # Filter for counties that have at least 50% of people engaged in public work
  filter(public_work>=50)
```

`rename()` verb & `select()` verb:
Notice the difference between select and rename. In select you need to name all the columns you want to keep along with renaming one or more of them. With rename, you can just pick one column whose name you want to change. The verb you choose will depend on your position within an analysis.\
```{r}
counties %>%
   select(state,county,population,unemployment_rate=unemployment)
counties %>% 
   select(state,county,population,unemployment)  %>%
   rename(unemployment_rate=unemployment)
```
```{r}
#renaming a column after counting 
counties %>%
  # Count the number of counties in each state
  count(state) %>%
  # Rename the n column to num_counties
  rename(num_counties=n)
#renaming a column as part of a select
counties %>%
  # Select state, county, and poverty as poverty_rate
 select(state,county,poverty_rate=poverty)
```


















































