---
title: "R Programming Skill Track II"
author: "Wenxiao Zhou"
date: "5/28/2021"
output: html_document
---

# Conditionals and Control Flow

(1) & and |

Like relational operators, logical operators work perfectly fine with vectors and matrices.
```{r}
linkedin<-c(16,9,13,5,2,17,14)
facebook<-c(17,7,5,16,8,13,14)
views<-matrix(c(linkedin,facebook),nrow=2,byrow=TRUE)
```
```
Three questions:
When did LinkedIn views exceed 10 and did Facebook views fail to reach 10 for a particular day? Use the linkedin and facebook vectors.
When were one or both of your LinkedIn and Facebook profiles visited at least 12 times?
When is the views matrix equal to a number between 11 and 14, excluding 11 and including 14?
```
```{r}
# linkedin exceeds 10 but facebook below 10
linkedin>10 & facebook<10

# When were one or both visited at least 12 times?
linkedin>=12 | facebook>=12

# When is views between 11 (exclusive) and 14 (inclusive)?
views>11 & views<=14
```


(2) The If statement:
```
if (condition) {
  expr
}
```
Add an else statement:
```
if (condition) {
  expr1
} else {
  expr2
}
```
Else if statement:
```
if (condition1) {
  expr1
} else if (condition2) {
  expr2
} else if (condition3) {
  expr3
} else {
  expr4
}
```


# Loops

(1) While loop
```
while (condition) {
  expr
}
```

Stop the while loop: break

This seems like a great opportunity to include the break statement in the while loop you've been working on. Remember that the break statement is a control statement. When R encounters it, the while loop is abandoned completely.
```{r}
#Adapt the while loop such that it is abandoned when the speed of the vehicle is greater than 80. This time, the speed variable has been initialized to 88; keep it that way.

# Initialize the speed variable
speed <- 88
while (speed > 30) {
  print(paste("Your speed is", speed))
  # Break the while loop when speed exceeds 80
  if (speed>80) {
    break
  }
  if (speed > 48) {
    print("Slow down big time!")
    speed <- speed - 11
  } else {
    print("Slow down!")
    speed <- speed - 6
  }
}
```


Building a while loop from scratch:
```{r}
# Initialize i as 1 
i <- 1
# Code the while loop
while (i <= 10) {
  print(3*i)
  if (i%%8==0) {
    print(3*i)
  }
  i <- i + 1
}
```


(2) For loops

Two ways of loop over a vector: \
```
primes <- c(2, 3, 5, 7, 11, 13)

# loop version 1
for (p in primes) {
  print(p)
}

# loop version 2
for (i in 1:length(primes)) {
  print(primes[i])
}
```

loop over a list\
```
primes_list <- list(2, 3, 5, 7, 11, 13)

# loop version 1
for (p in primes_list) {
  print(p)
}

# loop version 2
for (i in 1:length(primes_list)) {
  print(primes_list[[i]])
}

Notice that you need double square brackets - [[ ]] - to select the list elements in loop version 2.
```

loop over a matrix \
```
In your workspace, there's a matrix ttt, that represents the status of a tic-tac-toe game. It contains the values "X", "O" and "NA". Print out ttt to get a closer look. On row 1 and column 1, there's "O", while on row 3 and column 2 there's "NA".
```
```
Instructions
Finish the nested for loops to go over the elements in ttt:

The outer loop should loop over the rows, with loop index i (use 1:nrow(ttt)).
The inner loop should loop over the columns, with loop index j (use 1:ncol(ttt)).
Inside the inner loop, make use of print() and paste() to print out information in the following format: "On row i and column j the board contains x", where x is the value on that position.
```
```{r}
ttt<-matrix(c("O",NA,"X",NA,"O","O","X",NA,"X"),nrow=3,byrow=TRUE)
# define the double for loop
for (i in 1:nrow(ttt)) {
  for (j in 1:ncol(ttt)) {
    print(paste("On row",i," and column ",j," the board contains ",ttt[i,j]))
  }
}
```


(3) Break & Next:
The `break` statement abandons the active loop: the remaining code in the loop is skipped and the loop is not iterated over anymore.
The `next` statement skips the remainder of the code in the loop, but continues the iteration.

```
Instructions:
Extend the for loop with two new, separate if tests as follows:

If the vector element's value exceeds 16, print out "This is ridiculous, I'm outta here!" and have R abandon the for loop (break).
If the value is lower than 5, print out "This is too embarrassing!" and fast-forward to the next iteration (next).
```
```{r}
# The linkedin vector has already been defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

# Adapt/extend the for loop
for (li in linkedin) {
  if (li > 10) {
    print("You're popular!")
  } else {
    print("Be more visible!")
  }
  # Add if statement with break
  if(li>16){
    print("This is ridiculous, I'm outta here!")
    break
  }
  # Add if statement with next
  if(li<5){
    print("This is too embarrassing!")
    next
  }
  print(li)
}
```


Building a for loop from scratch:

We already went ahead and defined a variable rquote. This variable has been split up into a vector that contains separate letters and has been stored in a vector chars with the `strsplit()` function.

Can you write code that counts the number of r's that come before the first u in rquote?

```{r}
# Pre-defined variables
rquote <- "r's internals are irrefutably intriguing"
chars <- strsplit(rquote, split = "")[[1]]

# Initialize rcount
rcount <- 0
# Finish the for loop
for (char in chars) {
  if(char=="r"){
      rcount<-rcount+1
  }
  if(char=="u"){
      break
  }
}
# Print out rcount
rcount
```


# Functions

Black Box Principle:

Input--Black Box--Output

Function documentation: `help(function_name)` or `?sd`

R matches function arguments by position or by name, and finally, some function arguments can have a default value, which can be overridden. If you do not specify the value of an argument that has no default, typically an error will occur.

```
Useful Trick \
`args() function` \
This is a function to learn about the arguments of a function without having to read through the entire documentation. 
```

(1) Introduction to Functions
```{r}
# Consult the documentation on the mean() function
help(mean)
# Inspect the arguments of the mean() function
args(mean)
```

```{r}
# mean(x,trim=0,na.rm=FALSE,...) as the default S3 method
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)
# Calculate the mean of the sum
avg_sum<-mean(linkedin+facebook) 
# Calculate the trimmed mean of the sum
avg_sum_trimmed<-mean(linkedin+facebook,trim=0.2)

# Inspect both new variables
avg_sum
avg_sum_trimmed

# The advanced usage of the mean function
# The linkedin and facebook vectors have already been created for you
linkedin1 <- c(16, 9, 13, 5, NA, 17, 14)
# Basic average of linkedin
mean(linkedin1)
# Advanced average of linkedin
mean(linkedin1,na.rm=TRUE)
```

Call `abs()` inside `mean()` to calculate the Mean Absolute Deviation


(2) Writing Functions
Solve a particular, well-defined problem; Black box principle; If it works, inner workings less important.
```
my_fun<-function(arg1,arg2){
       body
}
```

```{r}
# Define the function hello()
hello<-function(){
    print("Hi there!")
    return(TRUE)
}
# Call the function hello()
hello()
```
```{r}
# Define pow_two() function, with print_info argument equals true as default
pow_two <- function(x, print_info = TRUE) {
  y <- x ^ 2
  if (print_info) {
    print(paste(x, "to the power two equals", y))
  }
  return(y)
}
# Some calls of the pow_two() function
pow_two(5)
pow_two(5, FALSE)
pow_two(5, TRUE)
```


Function scoping:

Variables that are defined inside a function are not accessible outside that function.\
```
pow_two <- function(x) {
  y <- x ^ 2
  return(y)
}
pow_two(4)
y
x
```

```{r}
#Exercise: Define a function to return the view counts
# The linkedin and facebook vectors have already been created for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)
facebook <- c(17, 7, 5, 16, 8, 13, 14)

# The interpret() can be used inside interpret_all()
interpret <- function(num_views) {
  if (num_views > 15) {
    print("You're popular!")
    return(num_views)
  } else {
    print("Try to be more visible!")
    return(0)
  }
}

# Define the interpret_all() function
# views: vector with data to interpret
# return_sum: return total number of views on popular days?
interpret_all <- function(views, return_sum=TRUE) {
  count <- 0
  for (v in views) {
    count=count+interpret(v)
  }
  if (return_sum) {
  return(count)
  } else {
   return(NULL)
  }
}
# Call the interpret_all() function on both linkedin and facebook
interpret_all(linkedin)
interpret_all(facebook)
```


(3) R Packages:

Install package: `install.packages("package_name")`; CRAN: Comprehensive R Archive Network  

Load package: `search()` \
7 packages are attached by default \
`library(package_name)`

`require()` function is similar to `library()` function: it loads packages into your R session. The only difference appears when you're trying to oad a package that is not yet installed.

If you use `library()` function directly, you will get an error; while use `require()` function, you only get a warning message.

```{r}
# Check out the currently attached packages again
search()
```


Different ways to load a package

The library() and require() functions are not very picky when it comes down to argument types: both `library(rjson)` and `library("rjson")` work perfectly fine for loading a package.

**You have to load packages one at a time with `require()` and `library()`** \
```
Two ways to load a package:
# Chunk 1
library(data.table)
require(rjson)

# Chunk 2
library("data.table")
require(rjson)
```


# The apply family

(1) lapply: 
apply function over list or vector \
function can return R objects of different classes \
list necessary to store heterogeneous content \
always return a list \
```{r}
#NYC: lapply()
nyc<-list(pop=8405837,
          boroughs=c('Manhattan','Bronx','Brooklyn','Queens','Staten Island'),
          capital=FALSE)
lapply(nyc,class)
```
```{r}
#Cities: for 
cities<-c("New York","Paris","London","Tokyo","Rio de Janeiro","Cape Town")
lapply(cities,nchar)
```
```{r}
cities<-c("New York","Paris","London","Tokyo","Rio de Janeiro","Cape Town")
# unlist function help to return a vector
unlist(lapply(cities,nchar))
```

For customized function:
```{r}
oil_prices<-list(2.37,2.49,2.18,2,22,2.47,2.32)
multiple<-function(x,factor){
  factor*x
}
result<-lapply(oil_prices,multiple,factor=3)
str(result)
unlist(result)
```
```{r}
# Code from previous exercise:
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

# Write function select_first()
select_first <- function(x) {
  x[1]
}
# Apply select_first() over split_low: names
names<-lapply(split_low,select_first)
# Write function select_second()
select_second<-function(x){
  x[2]
}
# Apply select_second() over split_low: years
years<-lapply(split_low,select_second)
```

**anonymous function in R** 
Writing your own functions and then using them inside `lapply()`
```{r}
# Definition of split_low
pioneers <- c("GAUSS:1777", "BAYES:1702", "PASCAL:1623", "PEARSON:1857")
split <- strsplit(pioneers, split = ":")
split_low <- lapply(split, tolower)

# Generic select function
select_el <- function(x, index) {
  x[index]
}
# Use lapply() twice on split_low: names and years
names<-lapply(split_low,select_el,index=1)
years<-lapply(split_low,select_el,index=2)
```


(2) sapply: \
When apply `sapply()` function, the result is a named vector, which contains the same information as the vector we obtained earlier using `unlist()` and `lapply()` together.

```{r}
# Cities: sapply()
cities<-c("New York","Paris","London","Tokyo","Rio de Janeiro","Cape Town")
first_and_last<-function(name){
  name<-gsub(" ","",name)
  letters<-strsplit(name,split="")[[1]]
  c(first=min(letters),last=max(letters))
}
sapply(cities,first_and_last)
```

Comparison between `lapply()` function and `sapply()` function
```{r}
one<-c(3,7,9,6,-1)
two<-c(6,9,12,13,5)
three<-c(4,8,3,-1,-3)
four<-c(1,4,7,2,-2)
five<-c(5,7,9,4,2)
six<-c(-3,5,8,9,4)
seven<-c(3,6,9,4,1)
temp<-list(one,two,three,four,five,six,seven)
# Finish function definition of extremes_avg
extremes_avg <- function(x) {
  ( min(x) + max(x)) / 2
}
# Apply extremes_avg() over temp using sapply()
sapply(temp,extremes_avg)
# Apply extremes_avg() over temp using lapply()
lapply(temp,extremes_avg)
```
```{r}
# Apply `sapply()` funciton with return vectors
# Create a function that returns min and max of a vector: extremes
extremes <- function(x) {
  c(min = min(x), max = max(x))
}
# Apply extremes() over temp with sapply()
sapply(temp,extremes)
# Apply extremes() over temp with lapply()
lapply(temp,extremes)
```
```{r}
# In some cases, sapply can't simplify
# Definition of below_zero()
below_zero <- function(x) {
  return(x[x < 0])
}
# Apply below_zero over temp using sapply(): freezing_s
freezing_s<-sapply(temp,below_zero)
# Apply below_zero over temp using lapply(): freezing_l
freezing_l<-lapply(temp,below_zero)

# Are freezing_s and freezing_l identical?
#identical(): Test Objects for Exact Equality
identical(freezing_s,freezing_l)
```

**`sapply() doesn't simplify the list of `NULL's`. That's because the 'vector-version' of a list of NULL's would simply be a NULL, which is no longer a vector with the same length as the input.**


(3) vapply:
apply function over list or vector \
explicitly sppecify output format \
(However, when using vapply, you have to explicitly say what the type of the return value will be. In sapply, this is not required nor possible.)

Comparison of `sapply()` and `vapply()`
```{r}
cities<-c("New York","Paris","London","Tokyo","Rio de Janeiro","Capge Town")
sapply(cities,nchar)
#vapply(X,FUN,FUN.VALUE,...,USE.NAMES=TRUE)
vapply(cities,nchar,numeric(1))
```

using the numeric() function, by setting FUN.VALUE to numeric(1), which tells the `vapply()` function that `nchar()` should return a single numerical value. The result is exactly the same as the `sapply()` function from before. However, this 'pre-specification' of FUN's return value makes `vapply()` a safer alternative to `sapply()`.

**This little bit of extra work in defining the FUN.VALUES arguments has the benefit that you really have to think about what your function will return without blindly assuming that the `sapply()` function will handle every case for you! **

```{r}
#unique_letters()
unique_letters<-function(name){
  name<-gsub(" ","",name)
  letters<-strsplit(name,split="")[[1]]
  unique(letters)
}
cities<-c("New York","Paris","London","Tokyo","Rio de Janeiro","Capge Town")
sapply(cities,unique_letters)
#error in the below function
#vapply(cities,unique_letters,character(5))
```

```{r}
one<-c(3,7,9,6,-1)
two<-c(6,9,12,13,5)
three<-c(4,8,3,-1,-3)
four<-c(1,4,7,2,-2)
five<-c(5,7,9,4,2)
six<-c(-3,5,8,9,4)
seven<-c(3,6,9,4,1)
temp<-list(one,two,three,four,five,six,seven)
# Definition of basics()
basics <- function(x) {
  c(min = min(x), mean = mean(x), max = max(x))
}
vapply(temp,basics,numeric(3))
```

From sapply to vapply:

```{r}
# Convert to vapply() expression
sapply(temp, max)
vapply(temp, max,numeric(1))

# Convert to vapply() expression
sapply(temp, function(x, y) { mean(x) > y }, y = 5)
vapply(temp, function(x, y) { mean(x) > y }, y = 5,logical(1))
```



# Utilities

Useful functions: 

(1)mathematical: `abs()`,`round()`,`sum()`,`mean()`

(2)data structure: `seq()`,`rep()`,`sort()`
```{r}
rep(c(8,6,4,2),times=2)
rep(c(8,6,4,2),each=2)
```
`unlist()`: flatten (possibly embedded) lists to produce a vector
`str()`:display the structure of any R object

(3)`is.*()`,`as.*()`

(4)`append()`:concatenates the original vector and its reversed version to create a list of elements, double this vector's length
```{r}
#combination of seq() and rep()
rep(seq(1, 7, by = 2), times = 7)

li<-list(log=TRUE,
         ch="hello",
         int_vec=sort(rep(seq(8,2,by=-2),times=2)))
str(append(li,rev(li)))
str(rev(li))
```

```{r}
# The linkedin and facebook lists have already been created for you
linkedin <- list(16, 9, 13, 5, 2, 17, 14)
facebook <- list(17, 7, 5, 16, 8, 13, 14)
# Convert linkedin and facebook to a vector: li_vec and fb_vec
li_vec<-unlist(linkedin)
fb_vec<-unlist(facebook)
# Append fb_vec to li_vec: social_vec
social_vec<-append(li_vec,fb_vec)
# Sort social_vec
sort(social_vec,decreasing=TRUE)
```


Regular Expressions:

Sequence of (meta) characters \
Pattern existence \
Pattern replacement \
Pattern extraction \
`grep()`,`grepl()` \
`sub()`,`gsub()` \

```{r}
#grep, grepl, regexpr, gregexpr and regexec search for matches to argument pattern within each element of a character vector: they differ in the format of and amount of detail in the results.
animals<-c("cat","moose","impala","ant","kiwi")
#grepl(pattern=<regex>,x=<string>)
grepl(pattern="a",x=animals)
# ^ matches the empty string at the beginning of a line
grepl(pattern='^a',x=animals)
# $ matches the empty string at the end of a line
grepl(pattern='a$',x=animals)
```
```{r}
# For more regular expression
?regex
```


Comparison of `grepl()` and `grep()` functions: \
`grep()`: This function returns a vector of indices of the elements of x that yield a match.  \
Compare the `grepl` command, that gives a vector of logicals, with the `grep` command, that gives the indices of the TRUE elements that the grepl function returns

One way in which you could compare grep and grepl would be using the which() function

```{r}
animals<-c("cat","moose","impala","ant","kiwi")
which(grepl(pattern="a",x=animals))
grep(pattern="^a",x=animals)
```


`sub()`,`gsub()` functions:
```{r}
animals<-c("cat","moose","impala","ant","kiwi")
#sub(pattern=<regex>,replacement=<str>,x=<str>)
sub(pattern="a",replacement="o",x=animals)
gsub(pattern="a",replacement="o",x=animals)
```

**sub() function only looks for the first match in the string, and if it finds it, replaces it with the replacement argument, and immediately stops looking. If you want to replace every single match of a pattern in a string with the replacement argument, you should try the gsub function instead**

OR metacharacter or vertical bar: similar to the or operator for combining logicals.
```{r}
gsub(pattern="a|i",replacement="_",x=animals)
```

Summary:

`grepl()`, which returns TRUE when a pattern is found in the corresponding character string. \

`grep()`, which returns a vector of indices of the character strings that contains the pattern.
```{r}
# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "dalai.lama@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")
# Use grepl() to match for "edu"
#Use grepl() to generate a vector of logicals that indicates whether these email addressess contain "edu". Print the result to the output.
grepl(pattern="edu",x=emails)

# Use grep() to match for "edu", save result to hits
hits<-grep(pattern="edu",x=emails)
# Subset emails using hits
emails[hits]
```

```
You can use the caret, **^**, and the dollar sign, **$** to match the content located in the start and end of a string, respectively. This could take us one step closer to a correct pattern for matching only the ".edu" email addresses from our list of emails. But there's more that can be added to make the pattern more robust:

@, because a valid email must contain an at-sign.
.*, which matches any character (.) zero or more times (*). Both the dot and the asterisk are metacharacters. You can use them to match any character between the at-sign and the ".edu" portion of an email address.
\\.edu$, to match the ".edu" part of the email at the end of the string. The \\ part escapes the dot: it tells R that you want to use the . as an actual character.
```

```{r}
#follow up the previous example
# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "dalai.lama@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use grepl() to match for .edu addresses more robustly
grepl(pattern='.*@.*\\.edu$',x=emails)
# Use grep() to match for .edu addresses more robustly, save result to hits
hits=grep(pattern='.*@.*\\.edu$',x=emails)
# Subset emails using hits
emails[hits]

```


Since `grep()` and `grepl()` simply check whether a regular expression could be matched with a character vector, `sub()` and `gsub()` take one step further: \
you can specify a `replacement` argument
```{r}
# The emails vector has already been defined for you
emails <- c("john.doe@ivyleague.edu", "education@world.gov", "global@peace.org",
            "invalid.edu", "quant@bigdatacollege.edu", "cookie.monster@sesame.tv")

# Use sub() to convert the email domains to datacamp.
sub(pattern='@.*\\.edu$',replacement='@datacamp.edu',x=emails)
```

```
New regular expressions:

.*: A usual suspect! It can be read as "any character that is matched zero or more times".
\\s: Match a space. The "s" is normally a character, escaping it (\\) makes it a metacharacter.
[0-9]+: Match the numbers 0 to 9, at least once (+).
([0-9]+): The parentheses are used to make parts of the matching string available to define the replacement. 
The \\1 in the replacement argument of sub() gets set to the string that is captured by the regular expression [0-9]+.
```

```{r}
#Example
awards <- c("Won 1 Oscar.",
  "Won 1 Oscar. Another 9 wins & 24 nominations.",
  "1 win and 2 nominations.",
  "2 wins & 3 nominations.",
  "Nominated for 2 Golden Globes. 1 more win & 2 nominations.",
  "4 wins & 1 nomination.")

sub(".*\\s([0-9]+)\\snomination.*$", "\\1", awards)
```
*Can you explain why all of this happened? *\
*The ([0-9]+) selects the entire number that comes before the word “nomination” in the string, and the entire match gets replaced by this number because of the \\1 that reference to the content inside the parentheses.*


Times & Dates
```{r}
today<-Sys.Date()
today
class(today)
now<-Sys.time()
now
class(now)
#The important class here is POSIXct. Apart from providing rich functionality for calculus and formatting, this class makes sure that the dates and times in R are compatible across different operating systems, according to the POSIX standard.
```

```{r}
#create a date objects
#by default
my_date<-as.Date("1971-05-14")
my_date
#setting the date format
my_date<-as.Date("1971-14-05",format="%Y-%d-%m")
my_date
#create POSIXct objects
my_time<-as.POSIXct("1971-05-14 11:25:15")
```

Date can perform calculation.
```{r}
my_date2<-as.Date("1998-09-29")
my_date2-my_date

my_time2<-as.POSIXct('1974-07-14 21:11:55 CET')
my_time2-my_time
```

Under the hood:
A Date object is simply a more advanced representation of the number of days since the first of January in 1970. 
```{r}
my_date<-as.Date('1997-06-01')
unclass(my_date)
my_time<-as.POSIXct('1997-07-01')
unclass(my_time)
```

Dedicated R packages: `lubridate`, `zoo`, `xts`

Create and format dates:
```
The character string has to obey a format that can be defined using a set of symbols:
%Y: 4-digit year (1982)
%y: 2-digit year (82)
%m: 2-digit month (01)
%d: 2-digit day of the month (13)
%A: weekday (Wednesday)
%a: abbreviated weekday (Wed)
%B: month (January)
%b: abbreviated month (Jan)
```

```{r}
# Definition of character strings representing dates
str1 <- "May 23, '96"
str2 <- "2012-03-15"
str3 <- "30/January/2006"
# Convert the strings to dates: date1, date2, date3
date1 <- as.Date(str1, format = "%b %d, '%y")
date2<-as.Date(str2,format='%Y-%m-%d')
date3<-as.Date(str3,format='%d/%b/%Y')

#Convert dates to formatted strings
#Extract useful information from the dates as character strings using format(). From the first date, select the weekday. From the second date, select the day of the month. From the third date, you should select the abbreviated month and the 4-digit year, separated by a space.
format(date1, "%A")
format(date2,'%d')
format(date3,'%b %Y')
```


Create and format times:
```
Similar to working with dates, you can use as.POSIXct() to convert from a character string to a POSIXct object, and format() to convert from a POSIXct object to a character string.

Symbols:
%H: hours as a decimal number (00-23)
%I: hours as a decimal number (01-12)
%M: minutes as a decimal number
%S: seconds as a decimal number
%T: shorthand notation for the typical format %H:%M:%S
%p: AM/PM indicator

*For a full list of conversion symbols, consult the strptime documentation in the console: ?strptime*

*as.POSIXct() uses a default format to match character strings. In this case, it's %Y-%m-%d %H:%M:%S*
```

```{r}
# Definition of character strings representing times
str1 <- "May 23, '96 hours:23 minutes:01 seconds:45"
str2 <- "2012-3-12 14:23:08"

# Convert the strings to POSIXct objects: time1, time2
time1 <- as.POSIXct(str1, format = "%B %d, '%y hours:%H minutes:%M seconds:%S")
time2<-as.POSIXct(str2,format="%Y-%m-%d %H:%M:%S")

# Convert times to formatted strings
format(time1,'%M')
format(time2,'%I:%M %p')
```
```{r}
# Create vector pizza
pizza <- c("2021-06-05","2021-06-07","2021-06-12","2021-06-18","2021-06-23")
# Create differences between consecutive pizza days: day_diff
day_diff<-diff(pizza)
#average diff
mean(day_diff)
```

```{r}
astro<-c("20-Mar-2015","25-Jun-2015","23-Sep-2015","22-Dec-2015")
meteo<-c("March 1, 15","June 1, 15","September 1, 15","December 1, 15")
names(astro)<-c("spring","summer","fall","winter")
names(meteo)<-c("spring","summer","fall","winter")
# Convert astro to vector of Date objects: astro_dates
astro_dates<-as.Date(astro,format='%d-%b-%Y')
# Convert meteo to vector of Date objects: meteo_dates
meteo_dates<-as.Date(meteo,format="%B %d, %y")
# Calculate the maximum absolute difference between astro_dates and meteo_dates
print(max(abs(astro_dates-meteo_dates)))
```




