---
title: "R Programming Skill Track II"
author: "Wenxiao Zhou"
date: "5/28/2021"
output: html_document
---

# Conditionals and Control Flow

(1) & and |

Like relational operators, logical operators work perfectly fine with vectors and matrices.
```{r}
linkedin<-c(16,9,13,5,2,17,14)
facebook<-c(17,7,5,16,8,13,14)
views<-matrix(c(linkedin,facebook),nrow=2,byrow=TRUE)
```
```
Three questions:
When did LinkedIn views exceed 10 and did Facebook views fail to reach 10 for a particular day? Use the linkedin and facebook vectors.
When were one or both of your LinkedIn and Facebook profiles visited at least 12 times?
When is the views matrix equal to a number between 11 and 14, excluding 11 and including 14?
```
```{r}
# linkedin exceeds 10 but facebook below 10
linkedin>10 & facebook<10

# When were one or both visited at least 12 times?
linkedin>=12 | facebook>=12

# When is views between 11 (exclusive) and 14 (inclusive)?
views>11 & views<=14
```


(2) The If statement:
```
if (condition) {
  expr
}
```
Add an else statement:
```
if (condition) {
  expr1
} else {
  expr2
}
```
Else if statement:
```
if (condition1) {
  expr1
} else if (condition2) {
  expr2
} else if (condition3) {
  expr3
} else {
  expr4
}
```


# Loops

(1) While loop
```
while (condition) {
  expr
}
```

Stop the while loop: break

This seems like a great opportunity to include the break statement in the while loop you've been working on. Remember that the break statement is a control statement. When R encounters it, the while loop is abandoned completely.
```{r}
#Adapt the while loop such that it is abandoned when the speed of the vehicle is greater than 80. This time, the speed variable has been initialized to 88; keep it that way.

# Initialize the speed variable
speed <- 88
while (speed > 30) {
  print(paste("Your speed is", speed))
  # Break the while loop when speed exceeds 80
  if (speed>80) {
    break
  }
  if (speed > 48) {
    print("Slow down big time!")
    speed <- speed - 11
  } else {
    print("Slow down!")
    speed <- speed - 6
  }
}
```


Building a while loop from scratch:
```{r}
# Initialize i as 1 
i <- 1
# Code the while loop
while (i <= 10) {
  print(3*i)
  if (i%%8==0) {
    print(3*i)
  }
  i <- i + 1
}
```


(2) For loops

Two ways of loop over a vector: \
```
primes <- c(2, 3, 5, 7, 11, 13)

# loop version 1
for (p in primes) {
  print(p)
}

# loop version 2
for (i in 1:length(primes)) {
  print(primes[i])
}
```

loop over a list\
```
primes_list <- list(2, 3, 5, 7, 11, 13)

# loop version 1
for (p in primes_list) {
  print(p)
}

# loop version 2
for (i in 1:length(primes_list)) {
  print(primes_list[[i]])
}

Notice that you need double square brackets - [[ ]] - to select the list elements in loop version 2.
```

loop over a matrix \
```
In your workspace, there's a matrix ttt, that represents the status of a tic-tac-toe game. It contains the values "X", "O" and "NA". Print out ttt to get a closer look. On row 1 and column 1, there's "O", while on row 3 and column 2 there's "NA".
```
```
Instructions
Finish the nested for loops to go over the elements in ttt:

The outer loop should loop over the rows, with loop index i (use 1:nrow(ttt)).
The inner loop should loop over the columns, with loop index j (use 1:ncol(ttt)).
Inside the inner loop, make use of print() and paste() to print out information in the following format: "On row i and column j the board contains x", where x is the value on that position.
```
```{r}
ttt<-matrix(c("O",NA,"X",NA,"O","O","X",NA,"X"),nrow=3,byrow=TRUE)
# define the double for loop
for (i in 1:nrow(ttt)) {
  for (j in 1:ncol(ttt)) {
    print(paste("On row",i," and column ",j," the board contains ",ttt[i,j]))
  }
}
```


(3) Break & Next:
The `break` statement abandons the active loop: the remaining code in the loop is skipped and the loop is not iterated over anymore.
The `next` statement skips the remainder of the code in the loop, but continues the iteration.

```
Instructions:
Extend the for loop with two new, separate if tests as follows:

If the vector element's value exceeds 16, print out "This is ridiculous, I'm outta here!" and have R abandon the for loop (break).
If the value is lower than 5, print out "This is too embarrassing!" and fast-forward to the next iteration (next).
```
```{r}
# The linkedin vector has already been defined for you
linkedin <- c(16, 9, 13, 5, 2, 17, 14)

# Adapt/extend the for loop
for (li in linkedin) {
  if (li > 10) {
    print("You're popular!")
  } else {
    print("Be more visible!")
  }
  # Add if statement with break
  if(li>16){
    print("This is ridiculous, I'm outta here!")
    break
  }
  # Add if statement with next
  if(li<5){
    print("This is too embarrassing!")
    next
  }
  print(li)
}
```


Building a for loop from scratch:

We already went ahead and defined a variable rquote. This variable has been split up into a vector that contains separate letters and has been stored in a vector chars with the strsplit() function.

Can you write code that counts the number of r's that come before the first u in rquote?

```{r}
# Pre-defined variables
rquote <- "r's internals are irrefutably intriguing"
chars <- strsplit(rquote, split = "")[[1]]

# Initialize rcount
rcount <- 0
# Finish the for loop
for (char in chars) {
  if(char=="r"){
      rcount<-rcount+1
  }
  if(char=="u"){
      break
  }
}
# Print out rcount
rcount
```


# Functions








