---
title: "R Programming Skill Track III"
author: "Wenxiao Zhou"
date: "6/24/2021"
output: html_document
---

# The Art of Benchmarking

Premature optimization is the root of all evil. ----Popularized by *Donald Knuth*

One of the relatively easy optimization available is to use an up-to-date version of R. In general, R is very conservative, so upgrading doesn't break existing code. However, a new version will often provide free speed boosts for key functions.
```{r}
# Print the R version details using version
version
# Assign the variable major to the major component
major <- version$major
# Assign the variable minor to the minor component
minor <- version$minor
```

Benchmarking(基准化分析法)

1. We construct a function around the feature we wish to benchmark \
2. We time the function under different scenarios, e.g. data set

```
Timing with `system.time()`
user time is the CPU time charged for the execution of user instructions.
system time is the CPU time charged for execution by the system on behalf of the calling process.
elapsed time is approximately the sum of user and system, this is the number we typically care about
```

Microbenchmark Package: Compare functions(Each function is run multiple times)
```{r}
library("microbenchmark")
n<-1e8
microbenchmark(1:n,
               seq(n),
              # seq_by(n),
               time=10)  # Run each function 10 times
```

Comparing read time of CSV and RDS files:
```
# How long does it take to read movies from CSV?
system.time(read.csv("movies.csv"))
# How long does it take to read movies from RDS?
system.time(readRDS("movies.rds"))
```

**Operational differences: '<-' and '=': using the `<-` operator inside a function call will create a new(or overwrite an existing) object **

Elapsed time

Using `system.time()` is convenient, but does have drawbacks when comparing multiple function calls. Use `microbenchmark::microbenchmark()` function to solve this problem:

<The microbenchmark() function makes it easier to compare multiple function calls at once by compiling all the relevant information in a data frame. It does this by running each function call multiple times, recording the time it took for the function to run each time, then computing summary statistics for each expression as you can see here.>

```
# Load the microbenchmark package
library(microbenchmark)
# Compare the two functions
compare <- microbenchmark(read.csv("movies.csv"), 
                          readRDS("movies.rds"), 
                          times = 10)

# Print compare
compare
```

**Measuring the performance of your computer is called benchmarking, and you can do that with the benchmarkme package.**


The benchmarkme package \
The main function within this package is `benchmark_std()`. These benchmarks are standard R operations such as loops and matrix calculations. On a standard machine, this code will take around four minutes or so to run. Once the benchmark has completed, you can compare your results to other users with the plot function. This method generates a set of plots that allows you to compare your machine to theirs. The plots on the right are for the programming benchmark. This particular benchmark focuses on looping. 
My computer is highlighted by the vertical line and I'm ranked 141 out of 385 machines. However relatively speaking, the fastest machine is less than twice as fast. 

```{r}
# The benchmarkme package
#install.packages("benchmarkme")
library("benchmarkme")
# Run each benchmark 3 times
res<-benchmark_std(runs=3)
plot(res)
```

```{r}
# Load the benchmarkme package
library(benchmarkme)

# Assign the variable ram to the amount of RAM on this machine
# Obtain the amount of RAM on this machine using get_ram()
ram <- get_ram()
ram
# Assign the variable cpu to the cpu specs
#Obtain the CPU specifications using get_cpu().
cpu <- get_cpu()
cpu
```

The benchmarkme package allows you to run a set of standardized benchmarks and compare your results to other users. One set of benchmarks tests is reading and writing speeds.
```{r}
# res=benchmark_io(runs=1,size=5)
# records the length of time it takes to read and write a 5MB file.
#plot(res)
```



# Fine Tuning: Efficient Base R

General principle: memory allocation

In R, memory allocation happens automatically. R allocates memory in RAM to store variables. Try to minimize variable assignment for speed.

```
Example: sequence of integers 1,2,...,n
(1)The obvious and best way:
x<-1:n

(2)Not so bad:
x<-vector("numeric",n)  # length n
for (i in 1:n)
   x[i]<-i
   
(3)Don't ever do this:
x<-NULL #length zero
for(i in 1:n)
 x<-c(x,i)
```

**The first rule of R club: never, ever grow a vector**
```{r}
#Compare growing a vector and 
#(1) Slow code
growing <- function(n) {
    x <- NULL
    for(i in 1:n)
        x <- c(x, rnorm(1))
    x
}
# Use <- with system.time() to store the result as res_grow
n <- 30000
system.time(res_grow <-growing(n))

#(2) Faster code
# Fast code
pre_allocate <- function(n) {
    x <- numeric(n) # Pre-allocate
    for(i in 1:n) 
        x[i] <- rnorm(1)
    x
}
n <- 30000
system.time(res_allocate<-pre_allocate(n))
```


Importance of vectorizing your code: \
General rule: Calling an R function eventually leads to C or FORTRAN code. <This code is very heavily optimized>

Goal: Access the underlying C or FORTRAN code as quickly as possible. The fewer functions call the better.

*Many R functions are vectorized.* \
*Loop is slow: loop is one-million-calls to the function; vectorized is a single-call to the function.*

**The second rule of R club: use a vectorized solution wherever possible.**

```{r}
#Vectorized code: multiplication
x<-rnorm(10)
x2<-numeric(length(x))
for(i in 1:10)
  x2[i]<-x[i]*x[i]
#Rewrite that code using a vectorized solution. Hints:
#You should not use a for loop.
#The multiplication operator is vectorized!
x2_imp <-x*x
```
```{r}
#Vectorized code: calculating a log-sum
# x is a vector of probabilities
# Initial code
n <- 100
total <- 0
x <- runif(n)
for(i in 1:n) 
    total <- total + log(x[i])

# Rewrite in a single line. Store the result in log_sum
log_sum <-sum(log(x))
```


Data frames and matrices:

Key data structure in R is data frame. It can be copied in other languages, e.g. Python: pandas data frame

Matrices: It's a rectangular data structure, you can perform usual subsetting and extracting operations. But every element must be the same data type.

*Because all values in a matrix object must be the same data type, it is much faster to access the first column of a matrix than it is to access that of a data frame.*















# Diagnosing Problems: Code Profiling

# Turbo Charged Code: Parallel Programming











































































