---
title: "Introduction to the Tidyverse"
author: "Wenxiao Zhou"
date: "4/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Wrangling
```{r}
setwd("/Users/zhouwenxiao/Desktop/Study/R/Datacamp/Introduction to Tidyverse")
library(gapminder)
library(dplyr)
glimpse(gapminder)
str(gapminder)
```

The first step for analyzing the data is: `filter()` \
A pipe is a ``%>%`. It says "take whatever is before it, and feed it into the next step.
```{r}
# Filter the gapminder dataset for the year 1957
gapminder %>% 
  filter(year==1957)
# Filter for China in 2002
gapminder %>%
 filter(country=='China' , year==2002)
```

Second, sort data using `arrange()`\
```{r}
# Filter for the year 1957, then arrange in descending order of population
gapminder %>%
 filter(year==1957) %>%
 arrange(desc(pop))
```

Third, `mutate()` can change the variable in the dataset, or create new variable.\
```{r}
# Use mutate to change lifeExp to be in months, named it as lifeExpMonths
gapminder %>% mutate(lifeExpMonths=12*lifeExp)

#Now combine all three steps together to answer the following question:
#Find the countries with the highest life expectancy, in months, in the year 2007
# Filter, mutate, and arrange the gapminder dataset
gapminder %>%
  filter(year==2007)  %>%
  mutate(lifeExpMonths=12*lifeExp) %>%
  arrange(desc(lifeExpMonths))
```



## Data Visualization
```{r}
library(ggplot2)
gapminder_1952 <- gapminder %>%
  filter(year == 1952)
# a scatter plot with GDP per capita on the x-axis and life expectancy on the y-axis
ggplot(gapminder_1952, aes(x = pop, y = gdpPercap)) +
  geom_point()
```

If you plot gdpPerCap vs. lifeExp, you will see that higher income countries tend to have higher life expectancy.\ 
One problem with this plot, however, is that a lot of countries get crammed into the leftmost part of the x-axis. This is because the distribution of GDP per capita spans several orders of magnitude, with some countries in the tens of thousands of dollars and others in the hundreds. When one of your axes has that kind of distribution, it's useful to work with a `logarithmic scale` **that is, a scale where each fixed distance represents a multiplication of the value. **
```{r}
#log scale
gapminder_1952 <- gapminder %>%
  filter(year == 1952)
# Change this plot to put the x-axis on a log scale
ggplot(gapminder_1952, aes(x = pop, y = lifeExp)) +
  geom_point() +
  scale_x_log10()

#color
# Scatter plot comparing pop and lifeExp, with color representing continent
ggplot(gapminder_1952,aes(x=pop,y=lifeExp,color=continent)) +
  geom_point() +
  scale_x_log10()
#size
# Add the size aesthetic to represent a country's gdpPercap
ggplot(gapminder_1952, aes(x = pop, y = lifeExp, color = continent,size=gdpPercap)) +
  geom_point() +
  scale_x_log10()
```

Visualizing data by categorical variables by using `facet_wrap()`:\
ggplot2 lets you divide your plot into subplots to get one smaller graph for each categories, it's another powerful way to communicate relationships within your data. 
```{r}
# Scatter plot comparing pop and lifeExp, faceted by continent
ggplot(gapminder_1952,aes(x=pop,y=lifeExp)) +
  geom_point() +
  scale_x_log10() +
  facet_wrap(~continent)

#faceting by year
# Scatter plot comparing gdpPercap and lifeExp, with color representing continent
# and size representing population, faceted by year
ggplot(gapminder,aes(x=gdpPercap,y=lifeExp,color=continent,size=pop)) +
 geom_point() +
 scale_x_log10() +
 facet_wrap(~year)
```


## Grouping and Summarizing
```{r}
library(gapminder)
library(dplyr)
#summarize() function
# Summarize to find the median life expectancy
gapminder %>%
  summarize(medianLifeExp=median(lifeExp))
# Filter for 1957 then summarize the median life expectancy and the maximum GDP per capita
gapminder %>%
  filter(year==1957)  %>%
  summarize(medianLifeExp=median(lifeExp),
            maxGdpPercap=max(gdpPercap)) 
```

`group_by()`: summarizing data in groups
```{r}
# Find median life expectancy and maximum GDP per capita in each continent in 1957
gapminder  %>%
  filter(year==1957) %>%
  group_by(continent) %>%
  summarize(medianLifeExp=median(lifeExp),
            maxGdpPercap=max(gdpPercap))

#group by two variables combined
# Find median life expectancy and maximum GDP per capita in each continent/year combination
gapminder  %>%
  group_by(continent,year) %>%
  summarize(medianLifeExp=median(lifeExp),
            maxGdpPercap=max(gdpPercap))
```

Visualizing summarized data
```{r}
by_year <- gapminder %>%
  group_by(year) %>%
  summarize(medianLifeExp = median(lifeExp),
            maxGdpPercap = max(gdpPercap))

# Create a scatter plot showing the change in medianLifeExp over time
#Be sure to add expand_limits(y = 0) to make sure the plot's y-axis includes zero
ggplot(by_year,aes(x=year,y=medianLifeExp)) +
  geom_point() +
  expand_limits(y=0)
```
```{r}
# Summarize medianGdpPercap within each continent within each year: by_year_continent
by_year_continent<-gapminder  %>%
   group_by(continent,year) %>%
    summarize(medianGdpPercap=median(gdpPercap))

# Plot the change in medianGdpPercap in each continent over time
ggplot(by_year_continent,aes(x=year,y=medianGdpPercap,color=continent)) +
  geom_point() +
  expand_limits(y=0)
```
Another example
```{r}
# Summarize the median GDP and median life expectancy per continent in 2007
by_continent_2007<- gapminder %>%
  filter(year==2007) %>%
  group_by(continent) %>%
  summarize(medianLifeExp=median(lifeExp),
            medianGdpPercap=median(gdpPercap))

# Use a scatter plot to compare the median GDP and median life expectancy
ggplot(by_continent_2007,aes(x=medianGdpPercap,y=medianLifeExp,color=continent)) +
    geom_point()
```



##Types of Visualizations\
**bar plots**: good at comparing statistics for each of several categories.\
To create a bar plot, you use `geom_col()`, short for "column". There are two aesthetics in a bar plot. X is the categorical variable, Y is the variable that determines the height of the bars.

&NewLine;
**Notice that: the positions of `group_by` and `filter` don't affect the data wrangling results.**
```{r}
#Visualizing median GDP per capita by continent
# Summarize the median gdpPercap by continent in 1952
by_continent<-gapminder %>%
  group_by(continent) %>%
  filter(year==1952)  %>%
  summarize(medianGdpPercap=median(gdpPercap))

# Create a bar plot showing medianGdp by continent
ggplot(by_continent,aes(x=continent,y=medianGdpPercap)) +
  geom_col()
```
```{r}
#Another example
#Visualizing GDP per capita by country in Oceania
# Filter for observations in the Oceania continent in 1952
oceania_1952<-gapminder %>% filter(continent=="Oceania", year==1952)
oceania_1952
# Create a bar plot of gdpPercap by country
ggplot(oceania_1952,aes(x=country,y=gdpPercap)) +
 geom_col()
```

**line plots**: visualizing changes over time.\
This makes it clearer that what we care about is the upward or downward trend over time. The way you make a line plot is to change one part of the code. `geom_point` was the part of the code that specified you were making a scatter plot. Simply change that to `geom_line` to make a line plot.
```{r}
#Visualizing median GDP per capita over time
# Summarize the median gdpPercap by year, then save it as by_year
by_year<- gapminder %>%
  group_by(year) %>%
  summarize(medianGdpPercap=median(gdpPercap))

# Create a line plot showing the change in medianGdpPercap over time
ggplot(by_year,aes(x=year,y=medianGdpPercap)) +
  geom_line() +
  expand_limits(y=0)
```
```{r}
#Visualizing median GDP per capita by continent over time
# Summarize the median gdpPercap by year & continent, save as by_year_continent
by_year_continent<-gapminder %>%
 group_by(year,continent) %>%
 summarize(medianGdpPercap=median(gdpPercap))

# Create a line plot showing the change in medianGdpPercap by continent over time
ggplot(by_year_continent,aes(x=year,y=medianGdpPercap,color=continent)) +
 geom_line() +
 expand_limits(y=0)
```

**histograms**: which describe the distribution of a one-dimensional numeric variable. 
```{r}
#Visualizing the population
gapminder_1952 <- gapminder %>%
  filter(year == 1952) %>%
  mutate(pop_by_mil = pop / 1000000)

# Create a histogram of population (pop_by_mil)
ggplot(gapminder_1952,aes(x=pop_by_mil)) +
#set the number of bins to 50
  geom_histogram(bins=50)

#also can set "binwidth" parameter within the geom_histogram
```
```{r}
#log scale
gapminder_1952 <- gapminder %>%
  filter(year == 1952)
# Create a histogram of population (pop), with x on a log scale
ggplot(gapminder_1952,aes(x=pop)) +
 scale_x_log10()+
  geom_histogram()
```


**box plots**: which compare the distribution of a numeric variable among several categories.\
(A boxplot helps give more context to the shape of the earlier histogram)
```{r}
#Comparing GDP per capita across continents
gapminder_1952 <- gapminder %>%
  filter(year == 1952)
# Create a boxplot comparing gdpPercap among continents
p<-ggplot(gapminder_1952,aes(x=continent,y=gdpPercap)) +
 geom_boxplot() +
 scale_y_log10()

# Add a title to this graph
p+ggtitle("Comparing GDP per capitaacross continents")
```

