---
title: "R Notebook For R Markdown: The Definitive Guide"
output:
  html_document:
    df_print: paged
---

## Preface
This notebook is a self-guide book referencing "R  Markdown: The Definitive Guide", no commercial use, only for learning.

All the codes and resources are all from "https://bookdown.org/yihui/rmarkdown/", Any infringement does so at his own risk.

Notes: 

(1) Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

(2) Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

(3) When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 


## Chapter 1: Installation

Install the rmarkdown package in R:
```{r}
#install.packages('rmarkdown')
library(rmarkdown)
```

If you want to generate PDF output, you will need to install LaTeX. For R Markdown users who have not installed LaTeX before, we recommend that you install TinyTeX.
```{r}
install.packages('tinytex')
tinytex::install_tinytex()  # install TinyTeX
```












