---
title: "Importing & Cleaning Data with R Skill Track I"
author: "Wenxiao Zhou"
date: "5/28/2021"
output: html_document
---

# Importing data from flat files with utils
(1)read.csv()

print the structure of the dataset: `str()`
```{r}
setwd("/Users/zhouwenxiao/Desktop/Datacamp-R/Importing & Cleaning Data with R/Data for Introduction")
pools <- read.csv("swimming_pools.csv", stringsAsFactors = TRUE)
str(pools)
```

(2)read.delim()

```
Import the data in "hotdogs.txt" with read.delim(). Call the resulting data frame hotdogs. The variable names are not on the first line, so make sure to set the header argument appropriately.
Call summary() on hotdogs. This will print out some summary statistics about all variables in the data frame.
```
```{r}
# Path to the hotdogs.txt file: path
path <- file.path("/Users/zhouwenxiao/Desktop/Datacamp-R/Importing & Cleaning Data with R/Data for Introduction", "hotdogs.txt")
hotdogs <- read.table(path, 
                      sep = "\t", 
                      col.names = c("type", "calories", "sodium"))
head(hotdogs)
```

(3)Arguments:
```
Lily wants to have the one with the fewest calories while Tom wants to have the one with the most sodium.

Instructions:
Finish the read.delim() call to import the data in "hotdogs.txt". It's a tab-delimited file without names in the first row.
The code that selects the observation with the lowest calorie count and stores it in the variable lily is already available. It uses the function which.min(), that returns the index the smallest value in a vector.
Do a similar thing for Tom: select the observation with the most sodium and store it in tom. Use which.max() this time.
Finally, print both the observations lily and tom.
```
```{r}
# Finish the read.delim() call
hotdogs <- read.delim("/Users/zhouwenxiao/Desktop/Datacamp-R/Importing & Cleaning Data with R/Data for Introduction/hotdogs.txt", header = FALSE, col.names = c("type", "calories", "sodium"))
# Select the hot dog with the least calories: lily
lily <- hotdogs[which.min(hotdogs$calories), ]
# Select the observation with the most sodium: tom
tom <- hotdogs[which.max(hotdogs$sodium), ]
# Print lily and tom
lily
tom
```

(4)Column classes:

Specify the column types or column classes of the resulting data frame. You can do this by setting the colClasses argument to a vector of strings representing classes:
```
read.delim("my_file.txt", 
           colClasses = c("character",
                          "numeric",
                          "logical"))
```
```{r}
# Previous call to import hotdogs.txt
hotdogs <- read.delim("/Users/zhouwenxiao/Desktop/Datacamp-R/Importing & Cleaning Data with R/Data for Introduction/hotdogs.txt", header = FALSE, col.names = c("type", "calories", "sodium"))

# Display structure of hotdogs
str(hotdogs)

# Edit the colClasses argument to import the data correctly: hotdogs2
hotdogs2 <- read.delim("/Users/zhouwenxiao/Desktop/Datacamp-R/Importing & Cleaning Data with R/Data for Introduction/hotdogs.txt", header = FALSE, 
            col.names = c("type", "calories", "sodium"),
            colClasses =c("factor","NULL","numeric"))
# Display structure of hotdogs2
str(hotdogs2)
```

**read.table() is the main function \ read.csv()=wrapper for CSV \ read.delim()=wrapper for tab-delimited files**

read.csv, defaults: header=TRUE, sep=","
```
i.e. read.table("states.csv",header=TRUE,sep=",",stringsAsFactors=FALSE) equivalent to 
read.table("states.csv",stringsAsFactors=FALSE)
```
read.delim, defaults: header=TRUE, sep="\t"
```
i.e. read.table("states.txt",header=TRUE,sep='\t',stringsAsFactors=FALSE) equivalent to 
read.delim("states.txt",stringsAsFactors=FALSE)
```

If you have a look at the documentation of read-dot-table, read.csv2 and read.delim2 are two functions exist to deal with regional differences in representing numbers.
You'll notice that the states_nay use commas for decimal points, as opposed to the dot for states_aye-dot-csv. This means that they can't use the comma as the field-delimiter anymore, they need a semicolon.











