---
title: "Data Manipulation with R"
author: "Wenxiao Zhou"
date: "4/26/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Joining data with dplyr

**Joining tables**\
&NewLine;
```{r}
#Understanding the data
library(dplyr)
library(tidyr)
library(ggplot2)
help(glimpse) #within dplyr package
setwd("/Users/zhouwenxiao/Desktop/Data Manipulation with R")
parts<-readRDS("parts.rds")
part_categories<-readRDS("part_categories.rds")
```
```{r}
#Joining parts and part categories by id
#use suffix argument to add suffixes to variable names to replace the name.x and name.y
parts %>% 
	inner_join(part_categories, by = c("part_cat_id" = "id"), 
	    suffix=c("_part","_category"))
```

Joining with a one-to-many relationship\
**you'll also see that paying close attention to the number of rows before and after a join is an important part of understanding your data.**
```{r}
inventories<-readRDS("inventories.rds")
sets<-readRDS("sets.rds")
inventory_parts<-readRDS("inventory_parts.rds")
#Previously, we know `parts` dataset is a list of all LEGO parts, and a new table, `inventory_parts` has some additional information about these parts, such as the color_id of each part you would find in a specific LEGO kit.
# Combine the parts and inventory_parts tables
parts %>%
  inner_join(inventory_parts,by="part_num")
#the order for inner_join is reversed, either direction will yield a table that contains the same information.
```


Joining three or more tables\
You can string together multiple joins with `inner_join` and the pipe `(%>%)`
```{r}
sets %>%
	# Add inventories using an inner join 
	inner_join(inventories,by="set_num") %>%
	# Add inventory_parts using an inner join 
	inner_join(inventory_parts,by=c("id"="inventory_id"))
```
```{r}
#What is the most common color?
#Additional table: colors, which tell us the color of each part in each set
colors<-readRDS("colors.rds")
# Count the number of colors and sort
sets %>%
	inner_join(inventories, by = "set_num") %>%
	inner_join(inventory_parts, by = c("id" = "inventory_id")) %>%
	inner_join(colors, by = c("color_id" = "id"), suffix = c("_set", "_color")) %>%
	count(name_color) %>%
	top_n(1,n)

#Alternatively
sets %>%
	inner_join(inventories, by = "set_num") %>%
	inner_join(inventory_parts, by = c("id" = "inventory_id")) %>%
	inner_join(colors, by = c("color_id" = "id"), suffix = c("_set", "_color")) %>%
	count(name_color, sort = TRUE)
```

&NewLine;
&NewLine;

**Joining tables**\
```
#Left joining two sets by part and color
millennium_falcon <- inventory_parts_joined %>%
  filter(set_num == "7965-1")
star_destroyer <- inventory_parts_joined %>%
  filter(set_num == "75190-1")
millennium_falcon %>%
  left_join(star_destroyer, by = c("part_num", "color_id"), suffix = c("_falcon", "_star_destroyer"))
```
```
#Left joining two sets by color
# Aggregate Millennium Falcon for the total quantity in each part
millennium_falcon_colors <- millennium_falcon %>%
  group_by(color_id) %>%
  summarize(total_quantity = sum(quantity))

# Aggregate Star Destroyer for the total quantity in each part
star_destroyer_colors <- star_destroyer %>%
  group_by(color_id) %>%
  summarize(total_quantity = sum(quantity))

# Left join the Millennium Falcon colors to the Star Destroyer colors
millennium_falcon_colors %>%
  left_join(star_destroyer_colors,by="color_id",suffix=c("_falcon","_star_destroyer"))
```

Add a filter after the joining
```
#Finding an observation that doesn't have a match
inventory_version_1 <- inventories %>%
  filter(version == 1)
# Join versions to sets
sets %>%
  left_join(inventory_version_1,by="set_num") %>%
  # Filter for where version is na
  filter(is.na(version)==TRUE)
```

Right Joins:\
This shows us that left joins and right joins are mirror images of each other. 
```{r}
#Read data
parts<-readRDS("parts.rds")
part_categories<-readRDS("part_categories.rds")
#Counting part colors
parts %>%
	count(part_cat_id) %>%
	right_join(part_categories, by = c("part_cat_id" = "id")) %>%
	# Filter for NA
	filter(is.na(n)==TRUE)
```
```{r}
#In both left and right joins, there is the opportunity for there to be NA values in the resulting table. Fortunately, the replace_na function can turn those NAs into meaningful values.
parts %>%
	count(part_cat_id) %>%
	right_join(part_categories, by = c("part_cat_id" = "id")) %>%
	# Use replace_na to replace missing values in the n column
	replace_na(list(n=0))
```


Joining tables to themselves\
Child-parent table:
```
For example: themse tibble contains id, name, parent_id. If parent_id is NA, which means the id is the parent_id itself. Now, we can use self-join to get a table with all ids name and their parents' id and name correspondly.

themes %>% 
  inner_join(themes,by=c("parent_id"="id"),suffix=c("_child","_parent"))  %>%
  filter(name_child=="The Lord of the Rings")
```
```{r}
themes<-readRDS("themes.rds")
themes %>% 
	# Inner join the themes table
	inner_join(themes,by=c("id"="parent_id"),suffix=c("_parent","_child"))  %>%
	# Filter for the "Harry Potter" parent name 
	filter(name_parent=="Harry Potter") 
```

**When performing the self-join, notice the position of by variables may change the meaning of the parent and children positions, make sure what you are focusing to target.**\

```{r}
#Joining themes to their grandchildren
#We can go a step further than looking at themes and their children. Some themes actually have grandchildren: their children's children.

#Here, we can inner join themes to a filtered version of itself again to establish a connection between our last join's children and their children.
# Join themes to itself again to find the grandchild relationships
themes %>% 
  inner_join(themes, by = c("id" = "parent_id"), suffix = c("_parent", "_child")) %>%
  inner_join(themes,by=c("id_child"="parent_id"),suffix=c("_parent","_grandchild"))
```

```{r}
#Left-joining a table to itself
#So far, you've been inner joining a table to itself in order to find the children of themes like "Harry Potter" or "The Lord of the Rings".

#But some themes might not have any children at all, which means they won't be included in the inner join. As you've learned in this chapter, you can identify those with a left_join and a filter()
themes %>% 
  # Left join the themes table to its own children
  left_join(themes,by=c("id"="parent_id"),suffix=c("_parent","_child")) %>%
  # Filter for themes that have no child themes
  filter(is.na(id_child)==TRUE)
```


**Full, Semi, and Anti Joins**\
The full join: keep both results in left and right joins.
```
batman_parts <- batman %>%
  count(part_num, color_id, wt = quantity)

star_wars_parts <- star_wars %>%
  count(part_num, color_id, wt = quantity)
batman_parts %>%
  # Combine the star_wars_parts table 
   full_join(star_wars_parts,by=c("part_num","color_id"),suffix=c('_batman','_star_wars')) %>%
  # Replace NAs with 0s in the n_batman and n_star_wars columns 
    replace_na(list(n_batman=0,
                    n_star_wars=0))
```

```
Comparing Batman and Star Wars LEGO parts
#parts_joined includes the part number of each piece, the color id, and the number of each piece in the Star Wars and Batman themes. 

parts_joined <- batman_parts %>%
  full_join(star_wars_parts, by = c("part_num", "color_id"), suffix = c("_batman", "_star_wars")) %>%
  replace_na(list(n_batman = 0, n_star_wars = 0))
```

Filtering joins: keeps or removes observations from the first table; doesn't add new variables\
`semi_join()`: What observations in X are also in Y
`anti_join()`: What observations in X are not in Y

**A semi join differs from an inner join because an inner join will return one row of x for each matching row of y, where a semi join will never duplicate rows of x. This is a filtering join. We get a similar result as with inner_join() but the join result contains only the variables originally found in x = superheroes.**

```{r}
#What colors are included in at least one set?
#Besides comparing two sets directly, you could also use a filtering join like semi_join to find out which colors ever appear in any inventory part. Some of the colors could be optional, meaning they aren't included in any sets.
# Use inventory_parts to find colors included in at least one set
colors %>%
  semi_join(inventory_parts,by=c("id"="color_id"))
```
```{r}
#Which set is missing version 1
# Each set included in the LEGO data has an associated version number. We want to understand the version we are looking at to learn more about the parts that are included. Before doing that, we should confirm that there aren't any sets that are missing a particular version.

# Use filter() to extract version 1 
version_1_inventories <- inventories %>%
 filter(version==1)
# Use anti_join() to find which set is missing a version 1
sets %>% 
  anti_join(version_1_inventories,by="set_num")
```


**Visualize the sets differences**\
To compare two kinds of LEGO sets, we need to aggregate the data into separate themes, additionally, add a new column to understand what is the fractions of specific pieces that are part of each set.
```{r}
setwd("/Users/zhouwenxiao/Desktop/Data Manipulation with R")
inventories<-readRDS("inventories.rds")
inventory_parts<-readRDS("inventory_parts.rds")
sets<-readRDS("sets.rds")
themes<-readRDS("themes.rds")
colors<-readRDS("colors.rds")
inventory_parts_themes<-inventories %>%
  inner_join(inventory_parts,by=c("id"="inventory_id")) %>%
  arrange(desc(quantity))  %>%
  select(-id,-version)  %>%
  inner_join(sets,by="set_num")  %>%
  inner_join(themes,by=c("theme_id"="id"),
             suffix=c("_set","_theme"))
batman_colors <- inventory_parts_themes %>%
  # Filter the inventory_parts_themes table for the Batman theme
  filter(name_theme == "Batman") %>%
  group_by(color_id) %>%
  summarize(total = sum(quantity)) %>%
  # Add a fraction column of the total divided by the sum of the total 
  mutate(fraction=total/sum(total))

# Filter and aggregate the Star Wars set data; add a fraction column
star_wars_colors <- inventory_parts_themes %>%
  filter(name_theme=="Star Wars")  %>%
  group_by(color_id) %>%
  summarize(total=sum(quantity))  %>%
  mutate(fraction=total/sum(total))
```

&NewLine;
Combining sets\
```{r}
batman_colors %>%
  full_join(star_wars_colors, by = "color_id", suffix = c("_batman", "_star_wars")) %>%
  replace_na(list(total_batman = 0, total_star_wars = 0)) %>%
  inner_join(colors, by = c("color_id" = "id")) %>%
  # Create the difference and total columns
  mutate(difference = fraction_batman-fraction_star_wars,
         total = total_batman+total_star_wars) %>%
  # Filter for totals greater than 200
  filter(total>200)
```

&NewLine;
The last step is creating a barplot to show the differences of Batman and Star Wars.
```
#install.packages("forcats")
library(forcats)
help("fct_reorder")

colors_joined <- batman_colors %>%
  full_join(star_wars_colors, by = "color_id", suffix = c("_batman", "_star_wars")) %>%
  replace_na(list(total_batman = 0, total_star_wars = 0)) %>%
  inner_join(colors, by = c("color_id" = "id")) %>%
  mutate(difference = fraction_batman - fraction_star_wars,
         total = total_batman + total_star_wars) %>%
  filter(total >= 200) %>%
  mutate(name = fct_reorder(name, difference)) 

# Create a bar plot using colors_joined and the name and difference columns
ggplot(colors_joined, aes(name,difference, fill = name)) +
  geom_col() +
  coord_flip() +
  scale_fill_manual(values = color_palette, guide = "none") +
  labs(y = "Difference: Batman - Star Wars")
```



**Case Study: Stack Overflow Questions**\
Three of the Stack Overflow survey datasets are `questions`, `question_tags`,`tags`. Use `left_join` to keep all questions even those without a corresponding tag.
```{r}
questions<-readRDS("questions.rds")
question_tags<-readRDS("question_tags.rds")
tags<-readRDS("tags.rds")
questions_with_tags<-questions %>%
  left_join(question_tags, by = c("id" = "question_id")) %>%
  left_join(tags, by = c("tag_id" = "id")) %>%
  replace_na(list(tag_name="only-r"))
questions_with_tags %>% 
  # Group by tag_name
  group_by(tag_name) %>%
  # Get mean score and num_questions
  summarize(score = mean(score),
    #use n() function to get the total number of questions
        	num_questions = n()) %>%
  # Sort num_questions in descending order
  arrange(desc(num_questions))
```


Finding the gaps between questions and answers:
```{r}
answers<-readRDS("answers.rds")
questions %>%
  # Inner join questions and answers with proper suffixes
  inner_join(answers,by=c("id"="question_id"),suffix=c("_question","_answer")) %>%
  # Subtract creation_date_question from creation_date_answer to create gap
  mutate(gap = as.integer(creation_date_answer-creation_date_question))
```
```{r}
#Joining question and answer counts
# Count and sort the question id column in the answers table
answer_counts <- answers %>%
  count(question_id, sort = TRUE)

# Combine the answer_counts and questions tables
questions %>%
  left_join(answer_counts, by = c("id" = "question_id")) %>%
  # Replace the NAs in the n column
  replace_na(list(n = 0))
```

Joining questions, answers, and tags:
This will allow us to do a better job of identifying which R topics get the most traction on Stack Overflow. 
```{r}
answer_counts <- answers %>%
    count(question_id, sort = TRUE)
question_answer_counts <- questions %>%
    left_join(answer_counts, by = c("id" = "question_id")) %>%
    replace_na(list(n = 0))
question_answer_counts %>%
  # Join the question_tags tables
  inner_join(question_tags,by=c("id"="question_id")) %>%
  # Join the tags table
  inner_join(tags,by=c("tag_id"="id"))

#Determine on average, how many answers each questions gets
tagged_answers <- question_answer_counts %>%
    inner_join(question_tags, by = c("id" = "question_id")) %>%
    inner_join(tags, by = c("tag_id" = "id"))
tagged_answers %>%
  # Aggregate by tag_name
  group_by(tag_name)  %>%
  # Summarize questions and average_answers
  summarize(questions = n(),
            average_answers =mean(n)) %>%
  # Sort the questions in descending order
  arrange(desc(questions)) 
```


**bind_rows** verb:\
```{r}
library(lubridate)
questions_with_tags <- questions %>%
  inner_join(question_tags, by = c("id" = "question_id")) %>%
  inner_join(tags, by = c("tag_id" = "id"))
answers_with_tags <- answers %>%
  inner_join(question_tags, by = "question_id") %>%
  inner_join(tags, by = c("tag_id" = "id"))
# Combine the two tables into posts_with_tags
posts_with_tags <- full_join(questions_with_tags %>% mutate(type = "question"),
                              answers_with_tags %>% mutate(type = "answer"))

# Add a year column, then count by type, year, and tag_name
posts_with_tags %>%
  mutate(year=year(creation_date)) %>%
  count(type,year,tag_name)
```
```{r}
# Filter for the dplyr and ggplot2 tag names 
by_type_year_tag <- posts_with_tags %>%
  mutate(year = year(creation_date)) %>%
  count(type, year, tag_name)
by_type_year_tag_filtered <- by_type_year_tag %>%
  filter(tag_name %in% c("dplyr","ggplot2"))

# Create a line plot faceted by the tag name 
ggplot(by_type_year_tag_filtered, aes(year, n, color = type)) +
  geom_line() +
  facet_wrap(~ tag_name)
```


